---
title: "SuSiE initialize at LASSO problem"
author: "Yuxin Zou"
date: "02/16/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

```{r}
library(susieR)
library(glmnet)
```
Here are two example data that show the probelm of initialize SuSiE using lasso results.

There is one signal in these examples.

## Converge slowly

```{r}
dat = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_2.rds')
sim = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_2_sim_gaussian_1.rds')
sumstat = susieR:::univariate_regression(dat$X_sample, sim$Y)
R = cor(dat$X_sample)
```

```{r}
z_scores <- sumstat$betahat / sumstat$sebetahat
susie_plot(z_scores, y = "z", b=sim$meta$true_coef)
```

The underlying causal variable is `r which(sim$meta$true_coef!=0)`. The effects are `r sim$meta$true_coef[sim$meta$true_coef!=0]`. The z scores are `r z_scores[sim$meta$true_coef!=0]`.

SuSiE model (estimate prior variance, not estimate residual variance):
```{r}
mod = susie(dat$X_sample, sim$Y[,1], estimate_residual_variance = F, estimate_prior_variance=T,null_weight=0,
                               coverage=0.95,min_abs_corr=0.5,
                               tol=1e-3, s_init=list())
```

The objective is `r mod$elbo[mod$niter]`. The CS is
```{r}
summary(mod)$cs
```

```{r}
susie_plot(mod, y='PIP', b = sim$meta$true_coef)
```

If we initialize using LASSO result, the model converges after 369 iterations.

```{r}
fit.lasso = glmnet(dat$X_sample, sim$Y, family="gaussian", alpha=1, dfmax = 10)
lasso.b = fit.lasso$beta[,max(which(fit.lasso$df <= 10))]
s_init = susie_init_coef(which(lasso.b!=0), sumstat$betahat[lasso.b!=0], nrow(sim$meta$true_coef))
mod_init_lasso = susie(dat$X_sample, sim$Y, estimate_residual_variance = F, estimate_prior_variance=T, s_init = s_init, max_iter = 500)
```

The objective is `r mod_init_lasso$elbo[mod_init_lasso$niter]`. The CS is
```{r}
summary(mod_init_lasso)$cs
```

```{r}
susie_plot(mod_init_lasso, y='PIP', b = sim$meta$true_coef)
```

## False Positive

```{r}
dat = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_20.rds')
sim = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_20_sim_gaussian_1.rds')
sumstat = susieR:::univariate_regression(dat$X_sample, sim$Y)
R = cor(dat$X_sample)
```

```{r}
z_scores <- sumstat$betahat / sumstat$sebetahat
susie_plot(z_scores, y = "z", b=sim$meta$true_coef)
```

The underlying causal variable is `r which(sim$meta$true_coef!=0)`. The effects are `r sim$meta$true_coef[sim$meta$true_coef!=0]`. The z scores are `r z_scores[sim$meta$true_coef!=0]`.

SuSiE model (estimate prior variance, not estimate residual variance):
```{r}
mod = susie(dat$X_sample, sim$Y[,1], estimate_residual_variance = F, estimate_prior_variance=T,null_weight=0,
                               coverage=0.95,min_abs_corr=0.5,
                               tol=1e-3, s_init=list())
```

The objective is `r mod$elbo[mod$niter]`. The CS is
```{r}
summary(mod)$cs
```

```{r}
susie_plot(mod, y='PIP', b = sim$meta$true_coef)
```

If we initialize using LASSO result, the model converges after 369 iterations.

```{r}
fit.lasso = glmnet(dat$X_sample, sim$Y, family="gaussian", alpha=1, dfmax = 10)
lasso.b = fit.lasso$beta[,max(which(fit.lasso$df <= 10))]
s_init = susie_init_coef(which(lasso.b!=0), sumstat$betahat[lasso.b!=0], nrow(sim$meta$true_coef))
mod_init_lasso = susie(dat$X_sample, sim$Y, estimate_residual_variance = F, estimate_prior_variance=T, s_init = s_init, max_iter = 500)
```

The objective is `r mod_init_lasso$elbo[mod_init_lasso$niter]`. The CS is
```{r}
summary(mod_init_lasso)$cs
```

```{r}
susie_plot(mod_init_lasso, y='PIP', b = sim$meta$true_coef)
```


