---
title: "SuSiE problem"
author: "Yuxin Zou"
date: "02/16/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

```{r}
library(susieR)
library(glmnet)
```
Here are two example data that show probelms of SuSiE initialization.

There are 2 signals in these examples.

## Example 1

```{r}
dat = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_27.rds')
sim = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_27_sim_gaussian_2.rds')
sumstat = susieR:::univariate_regression(dat$X_sample, sim$Y)
R = cor(dat$X_sample)
```

```{r}
z_scores <- sumstat$betahat / sumstat$sebetahat
susie_plot(z_scores, y = "z", b=sim$meta$true_coef)
```

The underlying causal variables are `r which(sim$meta$true_coef!=0)`. The effects are `r sim$meta$true_coef[sim$meta$true_coef!=0]`. The z scores are `r z_scores[sim$meta$true_coef!=0]`.

The correlation between the strongest signal and causal variables are:
```{r}
R[c(616, which(sim$meta$true_coef!=0)), c(616, which(sim$meta$true_coef!=0))]
```

SuSiE model (estimate prior variance, not estimate residual variance):
```{r}
mod = susie(dat$X_sample, sim$Y[,1], estimate_residual_variance = F, estimate_prior_variance=T,null_weight=0,
                               coverage=0.95,min_abs_corr=0.5,
                               tol=1e-3, s_init=list())
```

The objective is `r round(mod$elbo[mod$niter],2)`. The CS is
```{r}
summary(mod)$cs
```

```{r}
susie_plot(mod, y='PIP', b = sim$meta$true_coef)
```

If we initialize using LASSO result, the model converges after **406** iterations.
```{r}
fit.lasso = glmnet(dat$X_sample, sim$Y, family="gaussian", alpha=1, dfmax = 10)
lasso.b = fit.lasso$beta[,max(which(fit.lasso$df <= 10))]
s_init = susie_init_coef(which(lasso.b!=0), sumstat$betahat[lasso.b!=0], nrow(sim$meta$true_coef))
start.t = proc.time()
mod_init_lasso = susie(dat$X_sample, sim$Y, estimate_residual_variance = F, estimate_prior_variance=T, s_init = s_init, max_iter = 500)
end.t = proc.time()
```

LASSO initialized at
```{r}
which(lasso.b!=0)
```

The model uses `r (end.t - start.t)['elapsed']` seconds. The objective is `r round(mod_init_lasso$elbo[mod_init_lasso$niter], 2)`. The CS is
```{r}
summary(mod_init_lasso)$cs
```

```{r}
susie_plot(mod_init_lasso, y='PIP', b = sim$meta$true_coef)
```

## Example 2

```{r}
dat = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_158.rds')
sim = readRDS('data/susie_rss_ukb_default_2.8_problem/small_data_158_sim_gaussian_2.rds')
sumstat = susieR:::univariate_regression(dat$X_sample, sim$Y)
R = cor(dat$X_sample)
```

```{r}
z_scores <- sumstat$betahat / sumstat$sebetahat
susie_plot(z_scores, y = "z", b=sim$meta$true_coef)
```

The underlying causal variable is `r which(sim$meta$true_coef!=0)`. The effects are `r sim$meta$true_coef[sim$meta$true_coef!=0]`. The z scores are `r z_scores[sim$meta$true_coef!=0]`.

SuSiE model (estimate prior variance, not estimate residual variance):
```{r}
mod = susie(dat$X_sample, sim$Y[,1], estimate_residual_variance = F, estimate_prior_variance=T,null_weight=0,
                               coverage=0.95,min_abs_corr=0.5,
                               tol=1e-3, s_init=list())
```

The objective is `r round(mod$elbo[mod$niter],2)`. The CS is
```{r}
summary(mod)$cs
```

```{r}
susie_plot(mod, y='PIP', b = sim$meta$true_coef)
```

The correlation between the false positive CS and causal variables are:
```{r}
R[c(255, which(sim$meta$true_coef!=0)), c(255, which(sim$meta$true_coef!=0))]
```

If we initialize using LASSO result,

```{r}
fit.lasso = glmnet(dat$X_sample, sim$Y, family="gaussian", alpha=1, dfmax = 10)
lasso.b = fit.lasso$beta[,max(which(fit.lasso$df <= 10))]
s_init = susie_init_coef(which(lasso.b!=0), sumstat$betahat[lasso.b!=0], nrow(sim$meta$true_coef))
mod_init_lasso = susie(dat$X_sample, sim$Y, estimate_residual_variance = F, estimate_prior_variance=T, s_init = s_init, max_iter = 500)
```

LASSO initialized at
```{r}
which(lasso.b!=0)
```

The objective is `r round(mod_init_lasso$elbo[mod_init_lasso$niter], 2)`. The CS is
```{r}
summary(mod_init_lasso)$cs
```

```{r}
susie_plot(mod_init_lasso, y='PIP', b = sim$meta$true_coef)
```


